### 2019-10-21 paper/main.tex
Applied Ergonomics submission

% \documentclass[preprint,12pt]{elsarticle}
\documentclass[preprint,review,12pt]{elsarticle}
% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
% \documentclass[final,5p,times]{elsarticle}
% \documentclass[final,5p,times,twocolumn]{elsarticle}

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{lineno}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{natbib}
\usepackage{float}

\journal{Applied Ergonomics}

\begin{document}

    % TODO
    %  - cover letter
    %  - fig

    \begin{frontmatter}

        \title{Sex differences in upper limb musculoskeletal biomechanics during a lifting task}

        %% use the tnoteref command within \title for footnotes;
        %% use the tnotetext command for the associated footnote;
        %% use the fnref command within \author or \address for footnotes;
        %% use the fntext command for the associated footnote;
        %% use the corref command within \author for corresponding author footnotes;
        %% use the cortext command for the associated footnote;
        %% use the ead command for the email address,
        %% and the form \ead[url] for the home page:
        %%
        %% \title{Title\tnoteref{label1}}
        %% \tnotetext[label1]{}
        %% \author{Name\corref{cor1}\fnref{label2}}
        %% \ead{email address}
        %% \ead[url]{home page}
        %% \fntext[label2]{}
        %% \cortext[cor1]{}
        %% \address{Address\fnref{label3}}
        %% \fntext[label3]{}

        \author[1]{Martinez Romain\corref{cor1}}
        \ead{martinez.staps@gmail.com}
        \author[1]{Assila Najoua}
        \author[1]{Goubault Etienne}
        \author[1]{Begon Mickaël}

        \address[1]{School of Kinesiology and Exercise Science, Faculty of Medicine, University of Montreal}

        \cortext[cor1]{corresponding author.}

        \begin{abstract}
            Women experience higher prevalence of work-related upper limb musculoskeletal disorders compared to men.
            Previous studies have investigated the biological, kinematic and electromyographic sex-related differences during a lifting task but the actual differences in musculoskeletal loads remain unknown.
            We investigated the sex differences in three musculoskeletal indicators: the sum of muscle activations, the sum of muscle forces and the relative time spent beyond a shear-compression dislocation ratio.
            A musculoskeletal model was scaled on 20 women and 20 men lifting a 6 or 12~kg box from hip to eye level.
            Women generated more muscle forces and activations than men, regardless of the lifted mass.
            Those differences occurred when the box was above shoulder level.
            In addition, women spent more time beyond a shear-compression dislocation ratio.
            Our work suggests higher musculoskeletal loads among women compared to men during a lifting task, which could be the result of poor technique and strength difference.
        \end{abstract}

        \begin{keyword}
            sex differences \sep upper limb \sep musculoskeletal modeling
        \end{keyword}

    \end{frontmatter}

    \section{Highlights}\label{sec:highlights}
    \begin{itemize}
        \item We investigated sex differences in upper limb musculoskeletal biomechanics during a lifting task.
        \item Women generated higher muscle forces and muscle activations when the box was above shoulder level and spent more time beyond a shear-compression dislocation ratio.
        \item Those musculoskeletal indicators can be used to evaluate work technique and estimate the underlying musculoskeletal loads.
    \end{itemize}

    \section{Introduction}\label{sec:introduction}
    \input{parts/1-introduction}

    \section{Materials and methods}\label{sec:materials-and-methods}
    \input{parts/2-materials-and-methods}

    \section{Results}\label{sec:results}
    \input{parts/3-results}

    \section{Discussion}\label{sec:discussion}
    \input{parts/4-discussion}

    \section{Acknowledgements}\label{sec:acknowledgements}

    The study was financially supported by the Institut de Recherche Robert-Sauvé en Santé et Sécurité du Travail [\# 2014--0045].
    The authors would like to thank the workers who participated as well as Marjolaine Corbeil, Landry Desmoulins and Damien Leflem for their contribution during the data collection and data labelling.

    \appendix
    \input{parts/5-appendix}

    \bibliographystyle{elsarticle-num-names}
    \bibliography{ref}

\end{document}